---
title: "Data Structure for Stan Model"
author: "Morgan Kain"
date: "10/19/2021"
output: html_document
---

### Example data structure from simulated data -- required data structure to run the stan model

##### It will likely be best to:

1. Create the data frames ind_occ_phi and ind_occ_p (see details below) and then take summaries of those for many of the indices


##### Dimensional indexes

```{r, echo = TRUE}

   ## n_pop     --- number of distinct sampling regions
     n_pop

   ## n_periods --- number of years (or seasons) but probably years in this model given one on-season per year
     periods[1, ]

   ## n_ind     --- number of unique individuals ever captured across all populations  
     all_ind.all 

   ## times_within --- time period in weeks of the bd on-season (period in which animals are congregating and exposed to bd)      
     times[1, ] 

   ## n_times   --- times_within * n_periods
     times[1, ] * periods[1, ]
  
   ## ind_occ   --- total number of catch opportunities (number of sampling events * number of unique individuals) 
   ##               calculated within each population and then summed across populations  
     ind_occ.all

   ## ind_occ_min1 --- same as ind_occ but using (number of sampling events - 1)   
     ind_occ_min1.all

```

##### Short index vectors

```{r, echo = TRUE}

  ## time            --- index designating week within season -- 1:times_within repeated for the number of periods
    rep(seq(times[1, ]), periods[1, ])

  ## time_per_period --- 1:n_times, arranged in a matrix, nrow = times_within, ncol = n_periods
    matrix(data = seq(times[1, ] * periods[1, ]), nrow = times, ncol = periods[1, ]) %>% head()
  
  ## periods         --- index that designates period for each value of "time"  
    one_pop$periods_time

  ## ind_occ_size    --- number of possible captures for each individual 
    ind_occ_size.all

  ## ind_occ_min1_size --- number of possible captures -1 for each individual   
    ind_occ_min1_size.all

  ## ind_in_pop        --- which population each individual belongs to    
    ind_in_pop.all

  ### all individuals' survival and capture probabilities are a vector, meaning that individuals are not designated in any meaningful way. 
   ### Thus, need an index vector for individuals for which entries correspond to unique individuals
    
  ## p_first_index     --- vector that designates which entries of the detection vector are the first entry for each unique individual
    p_first_index
    
  ## phi_first_index   --- vector that designates which entries of the survival vector are the first entry for each unique individual  
    phi_first_index

```

##### Long index vectors

These vectors are one of two different possible lengths (ind_occ, or ind_occ_min1). It is therefore easiest to construct two data frames and then pull out the columns as the individual vectors that are needed

```{r, echo = TRUE}

### for survival (phi) which uses everything but the last sampling event (as phi is survival to the next event)
head(ind_occ_phi.all, 20)

 ## ind_occ_min1_rep --- (ind in the above data frame) the individual associated with every entry of phi
 
 ## sampling_events_phi --- (sampling_events_phi) the sampling event associated with each possible capture of each individual
 
 ## offseason --- (offseason) a vector of 0s and 1s designating which of the sampling events is the last of a season

 ## pop_phi --- (pop) population associated with each entry

 ## ind_occ_rep --- (ind in the above data frame) the individual associated with every entry of p

 ## phi_zeros --- (phi_zeros) entries of phi that must be 0 because we caught an individual after this entry
 
### for detection, use every sampling event
head(ind_occ_p.all, 20)

 ## sampling_events_p --- (sampling_events_p) the sampling event associated with each possible capture of each individual
 
 ## periods_occ --- (periods_occ) the period for every entry

 ## pop_p --- (pop) population associated with each entry

 ## p_zeros --- (p_zeros) periods in which an individual was never captured (but in which the individual could have been present)

``` 

##### Covariates

```{r, echo = TRUE}

### Also easiest to construct a data frame of all of the bd measures and then grab columns of that data frame
head(X_bd.m.all, 20)

  ## N_bd --- (nrow in above data frame) number of bd measures (in all of the data) 

  ## X_bd --- (bd) the actual bd measured

  ## ii_bd --- (ind) the individual associated with each bd measure

  ## tt_bd --- (times) the times on which each bd was measured


### Covariates need to be in different structures depending on what part of the model they enter.

 ### For the bd submodel they must be in a matrix with rows = times and columns = populations

  ## temp --- temperature associated with each population
   temp

 ### For predictors on phi and p, they need to be a long vector associated with each entry
   
  ## time_gaps --- time between two consecutive sampling events
   ind_occ_phi.all$time_gaps[1:20]

```

##### Capture Data

```{r, echo = TRUE}

### Find it easiest to add the captures to the matrix from before on detection

  ## N_y --- number of detection opportunities 
   nrow(ind_occ_p.all)
  
  ## y --- 1s and 0s for captures or not
   ind_occ_p.all$captures[1:20]
  
### summaries of captures
   head(capture_range.all, 20)
   
  ## first --- (first) the first occasion on which each individual was captured

  ## last --- (final) the last occasion on which each individual was captured

  ## present --- matrix with rows n_ind and columns n_periods that designates the first and all times forward from when an individual was captured
   head(present.all, 20)

```

## Including Plots

```{r pressure, echo = FALSE}

```

